\subsection{Integration into KAGE} \label{methods:integration_into_kage}
Previously in section \ref{methods}, we described how we implemented GPU acceleration for three components of KAGE: the \textit{k}mer hashing, \textit{k}mer counting and genotyping components.
Several different methods were explored, and some yielded better performance than others.
Common for each GPU accelerated component is that they were GPU accelerated in isolation to keep their modular behaviour identical to the non-GPU accelerated versions.
Thus, each accelerated component could be used as drop-in replacements for their non-GPU counterparts in KAGE.

We chose the components accelerated using the methods yielding the best performance and used them as drop-in replacements in KAGE.
For each component we chose the following implementations:

\begin{table}[H]
\begin{center}
\begin{tabular}{lllll}
\multicolumn{1}{l|}{\textbf{Component}} & \multicolumn{1}{l}{\textbf{Method}} &  \\ \cline{1-2}
\multicolumn{1}{l|}{\textit{k}mer hashing} & \multicolumn{1}{l}{CuPy as a NumPy drop-in replacement} &  \\
\multicolumn{1}{l|}{\textit{k}mer counting} & \multicolumn{1}{l}{custom CUDA hash table} &  \\
\multicolumn{1}{l|}{genotyping (LOGPMF)} & \multicolumn{1}{l}{custom CUDA kernel} &  \\
\end{tabular}
\end{center}
\caption{
  The implementations chosen to be integrated into KAGE for GPU acceleration.
}
\label{methods:integration_into_kage:tables:chosen_methods}
\end{table}

Recall from section \ref{background:kage}, that KAGE is composed of two separate programs: \textit{kmer\_mapper} - performing the \textit{k}mer hashing and counting, and \textit{kage} - performing the genotype-probability computations supported by the \textit{k}mer counts provided by \textit{kmer\_mapper}.
We replaced KAGE's \textit{k}mer hashing and \textit{k}mer counting components in \textit{kmer\_mapper} with our GPU accelerated versions, and \textit{kage}'s LOGPMF function with our GPU accelerated version.
Then, we benchmarked each piece of software to ensure runtime speedup, despite our assessments made for each component previously.

After incorporating the GPU accelerated components into \textit{kmer\_mapper} and \textit{kage}, we benchmarked both programs which yielded the following runtimes:
\begin{table}[H]
\begin{center}
\begin{tabular}{lllll}
  \multicolumn{1}{l|}{\textbf{Program}} & \multicolumn{1}{l}{\textbf{CPU}} & \multicolumn{1}{l}{\textbf{GPU}} & \\ \cline{1-3}
  \multicolumn{1}{l|}{\textit{kmer\_mapper}} & \multicolumn{1}{l}{484 sec} & \multicolumn{1}{l}{61 sec} & \\
  \multicolumn{1}{l|}{\textit{kage}} & \multicolumn{1}{l}{68 sec} & \multicolumn{1}{l}{54 sec} & \\
\end{tabular}
\end{center}
\caption{
  Benchmarking results from integrating the GPU accelerated components into \textit{kmer\_mapper} and \textit{kage}.
  \textit{kmer\_mapper} and \textit{kage} together constitute the KAGE genotyping pipeline.
  This benchmark was ran on a high-end compute server with an AMD EPYC 7742 64-Core CPU and two Nvidia Tesla V100 GPUs.
  Both \textit{kmer\_mapper} and \textit{kage} were allowed 16 cores and the GPU versions were allowed one Nvidia Tesla V100 GPU.
}
\label{methods:integration_into_kage:tables:benchmark}
\end{table}

As can be seen in table \ref{methods:integration_into_kage:tables:benchmark}, the GPU accelerated \textit{k}mer hashing and counting components yielded significant runtime improvements to the \textit{k}mer counting portion (\textit{kmer\_mapper}) of KAGE.
GPU acceleration yielded faster runtimes for the genotype-probabilities computation portion (\textit{kage}) of KAGE, although a less dramatic improvement.
