\subsubsection{DNA and Chromosomes} \label{background:biology:dna_and_chromosomes}
\textit{Deoxyribonucleic acid} (DNA) is a type of molecule that carries the genetic instructions for the development, function, and reproduction of all known living organisms.
The molecule is composed of two connected strands that wind around each other into a double helix shape.
Each strand is made up of a sugar and phosphate backbone, with each sugar carrying one out of four bases, often referred to as \textit{nucleotides}. 
The four DNA nucleotides are: \textit{cytosine}, \textit{guanine}, \textit{adenine} and \textit{thymine}, and they are typically referred to as simply C, G, A and T respectively.
Furthermore, chemical bonds form between complementary pairs of nucleotides: AT/TA and CG/GC, meaning that A and T, and C and G are complementary nucleotides. 
The two strands are connected through the bonds formed between these nucleotides.
Since the two strands are sequences of complementary nucleotides, the two strands do in fact represent the same genetic information. 
In other words, one strand can be determined by reversing the other and exchanging each nuclotide by its complement.

DNA is organized into structures called chromosomes. 
Humans have 46 chromosomes made up by two sets of 23 chromosomes where each chromosome occurs twice. One is inherited from the male parent and the other is inherited from the female parent.

\begin{figure}[ht!]
\begin{center}
\scalebox{1}{
\begin{tikzpicture}

  \node at(2.25,2.5)(title){\textit{DNA}};
  \node at(-2,0)(title){\textit{strand 2}};
  \node at(-2,1.5)(title){\textit{strand 1}};

  % Nodes
  \node at(0,0)(n1){...};
  \node at(0.75,0)(n2){A};
  \node at(1.25,0)(n3){T};
  \node at(1.75, 0)(n4){G};
  \node at(2.25,0)(n5){G};
  \node at(2.75,0)(n5){C};
  \node at(3.25,0)(n5){T};
  \node at(3.75,0)(n5){G};
  \node at(4.5,0)(n5){...};

  \node at(0,1.5)(n1){...};
  \node at(0.75,1.5)(n2){T};
  \node at(1.25,1.5)(n3){A};
  \node at(1.75,1.5)(n4){C};
  \node at(2.25,1.5)(n5){C};
  \node at(2.75,1.5)(n5){G};
  \node at(3.25,1.5)(n5){A};
  \node at(3.75,1.5)(n5){C};
  \node at(4.5,1.5)(n5){...};

  % base pair connects
  \draw (0.75,0.5) -- (0.75,1);
  \draw (1.25,0.5) -- (1.25,1);
  \draw (1.75,0.5) -- (1.75,1);
  \draw (2.25,0.5) -- (2.25,1);
  \draw (2.75,0.5) -- (2.75,1);
  \draw (3.25,0.5) -- (3.25,1);
  \draw (3.75,0.5) -- (3.75,1);

\end{tikzpicture}
}
\caption{A conceptual representation of a DNA molecule made up of two strands. The strands are composed of nucleotides forming base pairs where A (adenine) and T (thymine), and C (cytosine) and G (guanine) are complements of each other.}
\label{background:biology:dna_and_chromosomes:figures:dna_strands}
\end{center}
\end{figure}
